---
title: "Analysis Of Categorical Data"
author: "Table of Content"

output:
  html_document:
    toc: true
    toc_depth: 5
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message=FALSE}
library(dplyr)
library(tidyr)
library(car)
library(knitr)
library(readr)
library(ResourceSelection)
library(ggplot2)
library(oddsratio)
```

** Uploading file to R Studio
```{r}
risk <- read_csv("~/R/ProjectGroup74_Data.csv")
head(risk)
```

** Checking structure of Dataset
```{r}
str(risk)
```

** Converting variables with binary response to Factor
```{r}
risk$anaemia <- factor(risk$anaemia, levels = c(0,1), labels = c("No","Yes"))
risk$diabetes <- factor(risk$diabetes, levels = c(0,1), labels = c("No","Yes"))
risk$high_blood_pressure <- factor(risk$high_blood_pressure, levels = c(0,1), labels = c("No","Yes"))
risk$sex <- factor(risk$sex, levels = c(0,1), labels = c("Female","Male"))
risk$smoking <- factor(risk$smoking, levels = c(0,1), labels = c("No","Yes"))
risk$DEATH_EVENT <- factor(risk$DEATH_EVENT,levels = c(0,1), labels = c("No","Yes"))

min(risk$age,na.rm = FALSE)
max(risk$age,na.rm = FALSE)

risk <- risk %>% mutate(Age_Group = 
                           case_when(age>=35 & age<=44 ~ '35 to 44 yrs',
                           age>=45 & age<=54 ~ '45 to 54 yrs',
                           age>=55 & age<=64 ~ '55 to 64 yrs',
                           age>=65 & age<=74 ~ '65 to 74 yrs',
                           age>=75 & age<=84 ~ '75 to 84 yrs',
                           age>=85 & age<=95 ~ '85 and above',))

risk$Age_Group <- factor(risk$Age_Group, levels = c('35 to 44 yrs','45 to 54 yrs','55 to 64 yrs','65 to 74 yrs','75 to 84 yrs','85 and above'),
                                 labels = c('35 to 44 yrs','45 to 54 yrs','55 to 64 yrs','65 to 74 yrs','75 to 84 yrs','85 and above'))

head(risk)
```
** Summary of Data
```{r}
mlr::summarizeColumns(risk) %>% knitr::kable(caption = "Summary of Dataset")
```
** Checking levels of Factor Vairables

```{r}
levels(risk$anaemia)
levels(risk$diabetes)
levels(risk$high_blood_pressure)
levels(risk$sex)
levels(risk$smoking)
levels(risk$DEATH_EVENT)
levels(risk$Age_Group)
```

# Statistical Modelling
## Model Fitting

```{r}

model_one<-glm(DEATH_EVENT~1,data=risk,family='binomial')
model_two<-glm(DEATH_EVENT~.,data=risk,family='binomial')
final_model <- step(model_one, 
                   scope = list(lower = model_one,
                                upper = model_two),
                   direction = "forward")

```
```{r}
summary(final_model)
final_model$coefficients
final_model$residuals
final_model$fitted.values
final_model$linear.predictors
final_model$deviance
final_model$aic
final_model$null.deviance
final_model$iter
final_model$df.residual
final_model$df.null
final_model

glm(formula = DEATH_EVENT ~ time + ejection_fraction + serum_creatinine + 
    age + serum_sodium, family = "binomial", data = risk)

```

** Final Model **
Logit(Probablity of Death by Cardiovascular Disease) = 9.49303 - 0.02089 x time - 0.07343 x ejection_fraction + 0.68599 x serum_creatinine + 0.04247 x age - 0.06456 x serum_sodium

## Residual Analysis
```{r}
summary(final_model)
par(mfrow = c(2,2))
plot(final_model)

plot(density(resid(final_model, type='response')))
lines(density(resid(final_model, type='response')), col='red')

plot(density(resid(final_model, type='pearson')))
lines(density(resid(final_model, type='pearson')), col='red')

plot(density(rstandard(final_model, type='pearson')))
lines(density(rstandard(final_model, type='pearson')), col='red')
```

In residuals vs fitted, the predicted values lie between -4 to 1 while beyond one are some outliers. We can see some linearity between values. In Normal Q-Q plot, the value are in the range between -3 to 3 with linearity between the values. In scale-location plot it is seen that the values lie between -4 to 2 while there are some outliers beyond 2. It is seen that, predicted values and sqr.root of std.deviance residuals intersect at 0 at x-axis and 1 at y-axis. In the residuals vs leverage plot, all values are located between 0.00 to 0.05 while there are some outliers beyond 0.05.


## Response Analysis
```{r}
summary(final_model)

survival_fracejec <- ggplot(data = risk, mapping = aes(x = DEATH_EVENT, y = ejection_fraction), color = "blue") +
  geom_bar(stat = "identity")+
  theme_light()+
  labs(x = "Survival Status")
survival_fracejec

survival_creatinine <- ggplot(data = risk, mapping = aes(x = DEATH_EVENT, y = serum_creatinine), color = "blue") +
  geom_bar(stat = "identity")+
  theme_light()+
  labs(x = "Survival Status")
survival_creatinine

survival_sodium <- ggplot(data = risk, mapping = aes(x = DEATH_EVENT, y = serum_sodium)) +
  geom_bar(stat = "identity")+
  theme_light()+
  labs(x = "Survival Status")
survival_sodium

```

## Goodness of Fit
```{r}

modelone_res <- model_one$deviance
modeltwo_res <- model_two$deviance
finalmodel_res <- final_model$deviance

modelone_dfres <- model_one$df.residual
modeltwo_dfres <- model_two$df.residual
finalmodel_dfres <- final_model$df.residual

modelone_resdf <- modelone_res/modelone_dfres
modeltwo_resdf <- modeltwo_res/modeltwo_dfres
finalmodel_resdf <- finalmodel_res/finalmodel_dfres

res <- c(modelone_res,modelone_dfres,modelone_resdf)

dfres <- c(modeltwo_res,modeltwo_dfres,modeltwo_resdf)

resdf <- c(finalmodel_res,finalmodel_dfres,finalmodel_resdf)

res
dfres
resdf

gender.table <- with(risk, table(DEATH_EVENT, sex)) 
gender.table
chisq.test(gender.table)

smoke.table <- with(risk, table(DEATH_EVENT, smoking)) 
smoke.table
chisq.test(smoke.table)

anaemia.table <- with(risk, table(DEATH_EVENT, anaemia)) 
anaemia.table
chisq.test(anaemia.table)

diabetes.table <- with(risk, table(DEATH_EVENT, diabetes)) 
diabetes.table
chisq.test(diabetes.table)

highbp.table <- with(risk, table(DEATH_EVENT, high_blood_pressure)) 
highbp.table
chisq.test(highbp.table)

highcreatinine.table <- with(risk, table(DEATH_EVENT, serum_creatinine)) 
highcreatinine.table
chisq.test(highcreatinine.table)

highejecfrac.table <- with(risk, table(DEATH_EVENT, ejection_fraction)) 
highejecfrac.table
chisq.test(highejecfrac.table)

highsodium.table <- with(risk, table(DEATH_EVENT, serum_sodium)) 
highsodium.table
chisq.test(highsodium.table)

```

## Confidence Intervals
```{r}
summary(final_model)

CInt <- exp(confint(final_model))
CI <- exp(confint.default(final_model))

CInt
CI


final.predict <- data.frame(DEATH_EVENT = 0, ejection_fraction = 60,time = 10,serum_creatinine = 1.90, age = 45, serum_sodium = 130)
alpha<-0.05
model.predict <- predict(object = final_model, newdata = final.predict, type = "response", se = TRUE)  
Interval<-model.predict$fit + qnorm(p = c(alpha/2, 1-alpha/2))*model.predict$se.fit  
Interval


```

## Hypothesis Test
```{r}
t.test(risk$age~risk$DEATH_EVENT)
t.test(risk$ejection_fraction~risk$DEATH_EVENT)
t.test(risk$serum_creatinine~risk$DEATH_EVENT)

summary(anova(final_model))
Anova(final_model , test = "LR")
```

## Sensitivity Analysis
```{r}
summary(final_model)

exp(cbind(Odds_and_OR=coef(final_model), confint(final_model)))
exp(coef(final_model))
exp(final_model$coefficients[2])

Anova(final_model)
```



# Reference
Dataset from GitHub by Dr. Vural Aksakalli. Reference of Website: vaksakalli/datasets. (2020). Retrieved 27 September 2020, from https://github.com/vaksakalli/datasets/blob/master/heart_failure.csv
